Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-06-24 kappa.tex
Describe the bedrock response model.

\subsection{Bedrock response}
\subsection{Surface mass balance}
\subsection{Atmospheric forcing}
\subsection{Numerical implementation}

>>>>

\subsection{Bedrock response}

The modelled ice sheet interacts with the underlying bedrock in two ways.
Firstly, bedrock temperatures are computed to a given depth by appending the
ice enthalpy model \eqn{enthalpy} with an underlying bedrock thermal model.
The only process accounted for is heat conduction, hence this model consists of
a simple diffusion equation.

\begin{equation}
    \pdv{T}{t} = \frac{k_b}{\rho_b c_b} \Delta T,
\end{equation}

where $k_b$, $\rho_b$ and $c_b$ are the thermal conductivity, density and
specific heat capacity of the bedrock. Bedrock temperature is conditioned at
depth by a fixed upwards geothermal heat flux $q_G$. The bedrock thermal model
is necessary for modelling the Cordilleran ice sheet on multi-millennial
time-scales, because temperature fluctuations caused by climate change and
isolating effects of the ice sheet penetrate several hundred meters into the
rock, and eventually affect basal ice temperatures.

Secondly, bedrock elevation evolve in response to the ice load. The bedrock
deformation model used in this thesis describes the flexure of an elastic
lithosphere on top of an infinite half-space viscous mantle
\citep{Lingle.Clark.1985}. It can be described by a single differential
equation of the bed elevation $b$,

\begin{equation}
    2\nu_m\,|\grad|\,\pdv{b}{t} + \rho_l g b + D\,\Delta^2 b = \sigma_{zzb},
\end{equation}

where $\nu_m$ is mantle viscosity, $\rho_l$ is lithosphere density, $D$ is the
lithosphere's flexural rigidity, and $\sigma_{zzb}$ corresponds to the ice load
\citep{Bueler.etal.2007}. Here, $\Delta^2$ designs the biharmonic operator of
linear elastic theory, while $|\grad|$ is a pseudo-differential operator
defined through the Fourier transform by \citet[Eqn.~6]{Bueler.etal.2007}. In
the left-hand part of this equation, the first term accounts for mantle
relaxation, the second for point-wise isostasy, and the third for elastic
flexure. Due to high mantle viscosity $\nu_m$, there exists a time lag between
the ice sheet growth and the isostatic bedrock response.
%$\nu=1\times10^{21}$\,Pa\,s
%$D=5\times10^{24}$\,N\,m
%$\rho_r = 3300$\,kg\,m$^{-3}$

\subsection{Surface mass balance}
\subsection{Atmospheric forcing}
\subsection{Numerical implementation}
